---
title: "Vegetation Monitoring Analysis"
author: "Manhatton"
date: "2/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install packages
install.packages('skimr', repos = "http://cran.us.r-project.org")
install.packages('kableExtra', repos = "http://cran.us.r-project.org")
install.packages('janitor', repos = "https://cran.r-project.org/package=janitor")
install.packages('pivottabler', repos = "http://cran.us.r-project.org")

## attach libraries
library(tidyverse)
library(readxl)
library(janitor)
library(dplyr)
library(here)
#library(skimr)  # not used so far
#library(kableExtra)  #not used so far
```

```{r DCrelevedata, include = FALSE}

### This is the code to read in the 2020 and 2021 data ---------------------------------

##to do:
## refactor function - low priority

##Find all releve xlsx within directory

file.list <- list.files(path = here::here("data", "2021 Format"), full.names = TRUE)

## Extract data from woody, forbs and grasses worksheets

 extract_data <- function(x)
{
woody <- read_xlsx(x, sheet = "Woody", skip = 8)%>%
  clean_names() %>%
  remove_empty(c("rows"))%>%
  mutate(rarity_status = as.character(rarity_status)) %>%
  select("scientific_name", "common_name", "x3" , "physiognamy", "mn_nativity", "mnwi", "rarity_status" = , "id" , "cc" , "cc_range", "x12" , "midpoint_cc", "p", "p_c")

forbs <- read_xlsx(x, sheet = "Forbs", skip = 8)%>%
 janitor::clean_names() %>%
 janitor::remove_empty(c("rows"))%>%
 mutate(rarity_status = as.character(rarity_status))  %>%
 select("scientific_name", "common_name", "x3" , "physiognamy", "mn_nativity", "mnwi", "rarity_status" = , "id" , "cc" , "cc_range", "x12" , "midpoint_cc", "p", "p_c")
  
grasses <- read_xlsx(x, sheet = "Grasses", skip = 8)%>%
  janitor::clean_names() %>%
  janitor::remove_empty(c("rows"))%>%
  mutate(rarity_status = as.character(rarity_status)) %>%
  select("scientific_name", "common_name", "x3" , "physiognamy", "mn_nativity", "mnwi", "rarity_status" = , "id" , "cc" , "cc_range", "x12" , "midpoint_cc", "p", "p_c")

relevecode <- read_xlsx(x, sheet = "Woody", range = "C2:C2", col_names = FALSE) #grabs the relevecode from C2 of Woody sheet

 data_frame <- (bind_rows(woody, forbs, grasses)) #combine the woody, forbs, and grasses data rows

 data_frame['releveCode'] <- list(relevecode[[1,1]]) #assign value from woody cell C2 to new column in combined dataframe (indicates releve code)
 
 ##getting the year out of the releve code
 n_last <- 4
 data_frame ['releveyear'] <- substr(relevecode, nchar(relevecode)-n_last + 1, nchar(relevecode) )
  
  return(data_frame)
}

 ##create empty dataframe with all the required columns
 
all_df <- data.frame("scientific_name"=character(), "common_name"=character(), "x3" =numeric(), "physiognamy"=character(), "mn_nativity" = character(), "mnwi"= character(), "rarity_status" = character(), "id" = character(), "cc" = numeric(), "cc_range" = character(), "x12" = numeric(), "midpoint_cc"= numeric(), "p" = numeric(), "p_c" = numeric(), releveCode = character(), releveyear = numeric()) 

##loop through all files in the folder 

for (x in file.list){
  print(x) #this helps troubleshoot if the code gets hung up somewhere
  single_df <- extract_data(x)
  all_df <- rbind(all_df, single_df)
}

##removes all rows with greater than 10 "NA" cells 

hi <- apply(all_df, 1, function(x) sum(is.na(x)))
all_df <- all_df[hi<10,]

## declutter the environment

rm(single_df)
rm(hi)
rm(x)

```

```{r DC_Old_Data, include = FALSE}

### This is the code to read in the processed data from 2019 and prior (just includes releve code, species and midpoint % cover)  -------

## Change this to the file for the 2019 and prior data that was compiled for R previously

old <- here("Data", "2019 and prior data", "Compiled Releve Data for R.xlsx")

  olddata <- read_excel(old, sheet = "Sheet1", col_name = TRUE) %>%
    select("Releve", "Species", "Midpoint % Cover") %>%
    janitor::clean_names() %>%
    add_column(releveyear = "")
  
##getting the year out of the releve code and creating column for it
 n_last <- 4
 olddata['releveyear'] <- substr(olddata$releve, nchar(olddata$releve)-n_last + 1, nchar(olddata$releve))
 
## declutter
 rm(old)
 rm(n_last)

```

```{r ReferenceSiteData, include=FALSE}

### This is the code to load in the GRG Reference Site Data ---------------------


file.ls <- list.files(here::here("data", "GRG Reference"), pattern = '\\.xlsx', full.names = TRUE)


ref_extract <- function(y)
{ 
  refdata <- read_xlsx(y, sheet= 1, skip = 1, col_names = TRUE) %>%
    janitor::remove_empty(c("rows"))%>%
    select("Genus", "Species", "3Freq" , "4Abund", "5Index")
    
  refcode <- read_xlsx(y, sheet = 1, range = "A1:A1", col_names = FALSE)
  refcode <- list(refcode[[1,1]])
  
  df<- bind_cols(refdata, refcode)
  
  return(df)
}

all_ref <- data.frame("genus" = character(), "species" = character(), "freq" = numeric() , "abund" = numeric(), "index"= numeric())

for (y in file.ls){
  print(y)
  single_ref <- ref_extract(y)
  all_ref <- rbind(all_ref, single_ref)
}

##removes all rows with greater than 4 "NA" cells 

empties <- apply(all_ref, 1, function(x) sum(is.na(x)))
all_ref <- all_ref[empties<4,]

##makes sure the abundance is numeric

all_ref$`4Abund` <-as.numeric(all_ref$`4Abund`)

## declutter
 rm(empties)
 rm(file.ls)
 rm(refdir)
 rm(y)
 rm(single_ref)

```

```{r AddCanopySubcanopyEtc, include=FALSE}
# 
# file.ls <- list.files(here::here("data", "GRG Reference"), pattern = '\\.xlsx', full.names = TRUE)
# 
# ref_extract2 <- function(y)
# { 
#   refdata2 <- read_xlsx(y, sheet= 1, skip = 1, col_names = TRUE) %>%
#     janitor::remove_empty(c("rows"))
#   
# ## find the row numbers where new sections start  
#   refdata2$row_num <- seq.int(nrow(refdata2))
#   
#   subcanopy <- which(grepl("Understory Trees", refdata2$Genus)) 
#   
#   shrubs <- which(grepl("Shrubs", refdata2$Genus)) 
#   
#   low_shrubs <- which(grepl("Low Shrubs", refdata2$Genus)) 
#   
#   vines <- which(grepl("Vines", refdata2$Genus)) 
#   
#   forbs <- which(grepl("Forbs", refdata2$Genus)) 
#   
#   grasses <- which(grepl("Grasses, Rushes and Sedges", refdata2$Genus)) 
#   
#   ferns <- which(grepl("Ferns and Fern Allies", refdata2$Genus))
#   
#   stop <- which(grepl("Exotic Invasive Species - Do Not Plant", refdata2$Genus))
#   
#   canopy <- refdata2 %>%
#     filter(row_num < subcanopy) %>%
#     dplyr::select(-row_num) %>%
#     add_column(HeightClass = "canopy")
#   subcanopy <- refdata2 %>%
#     slice(subcanopy, shrubs) %>%
#     dplyr::select(-row_num) %>%
#     add_column(HeightClass = "subcanopy")
#    shrubs <- refdata2 %>%
#     filter(row_num < shrubs > low_shrubs) %>%
#     dplyr::select(-row_num) %>%
#     add_column(HeightClass = "shrubs")
#    low_shrubs <- refdata2 %>%
#     slice(low_shrubs:vines) %>%
#     dplyr::select(-row_num) %>%
#     add_column(HeightClass = "low shrubs")
#    vines <- refdata2 %>%
#     slice(vines:forbs) %>%
#     dplyr::select(-row_num) %>%
#     add_column(HeightClass = "vines")
#    forbs <- refdata2 %>%
#     slice(forbs:grasses) %>%
#     dplyr::select(-row_num) %>%
#     add_column(HeightClass = "forbs")
#    grasses <- refdata2 %>%
#     slice(grasses:ferns) %>%
#     dplyr::select(-row_num) %>%
#     add_column(HeightClass = "graminoids")
#    ferns <- refdata2 %>%
#     slice(grasses:stop) %>%
#     dplyr::select(-row_num) %>%
#     add_column(HeightClass = "ferns")
#     
#   
#   subcanopy <-
#     
#   refcode2 <- read_xlsx(y, sheet = 1, range = "A1:A1", col_names = FALSE)
#   refcode2 <- list(refcode2[[1,1]])
#   
#   df<- bind_cols(refdata2, refcode)
#   
#   return(df)
# }
# 
# Ref_all <- data.frame("genus" = character(), "species" = character(), "freq" = numeric() , "abund" = numeric(), "index"= numeric())
# 
# for (y in file.ls){
#   print(y)
#   single_ref <- ref_extract(y)
#   all_ref <- rbind(all_ref, single_ref)
# }


```

```{r ordination_prep_table, include=FALSE}

##create new dataframe with just the scientific name, midpoint % cover, releve 

### For Reference:
### olddata = the 2019 and prior data prepped for analysis in excel
### all_df = the complete 2020 and 2021 releves
### ref_data = the GRG reference site data 

DCData1 <- all_df[ ,c("scientific_name", "midpoint_cc", "releveCode")] %>%
  group_by(scientific_name, releveCode) %>%
  slice(which.max(midpoint_cc)) %>%
  ungroup()

DCData2 <- olddata[ , c("species", "midpoint_percent_cover", "releve")] %>%
     dplyr::rename("scientific_name" = "species", 
                "midpoint_cc" = "midpoint_percent_cover", 
                "releveCode" = "releve")%>%
  group_by(scientific_name, releveCode) %>%
  slice(which.max(midpoint_cc)) %>%
  ungroup()

RefData <- all_ref[ , c("Genus", "Species", "4Abund", "...6")]
  
RefData <- unite(RefData, "scientific_name", Genus:Species, sep = " ", remove = TRUE) 

RefData <- dplyr::rename(RefData, "midpoint_cc" = "4Abund", 
                "releveCode" = "...6") %>%
  group_by(scientific_name, releveCode) %>%
  slice(which.max(midpoint_cc)) %>%
  ungroup()
  

AllData <-
  dplyr::bind_rows(
    #AllData, 
    DCData1, 
    DCData2, 
    RefData 
        )

```

```{r ordination_prep_pivot, include=FALSE}

## Create pivot table with species (row), releve (column) and midpoint percent cover (data)

dcveg <- AllData %>%
  mutate(midpoint_cc = ifelse(is.na(midpoint_cc), 0, midpoint_cc)) %>%
  pivot_wider(names_from = scientific_name, values_from = midpoint_cc, values_fn = list)

 
```

```{r DCA_OrdinationCode2019}

library(vegan)
library(MASS)
library(cluster.datasets)
library(ggplot2)

#dcveg <- read.csv("C:\\Users\\Meg\\OneDrive - University of Nebraska-Lincoln\\R practice\\DCVegMon\\Old R script\\RScript\\RelData.csv", row.names=1, header = TRUE, sep = ",")

#groups <- read.csv("N:/PARKS/Natural Resources/Monitoring/Transects, Releves, and Photo Monitoring/Archived Monitoring Data/Analyzed Releve Data/RScript/Groups.csv", row.names=1, header = TRUE, sep = ",")

 
dcveg[,-1] <- sapply(dcveg[,-1], function(x) as.numeric(as.character(x)))

dcveg[is.na(dcveg)] <- 0

dcveg <- column_to_rownames(dcveg, var="releveCode")

ord <- decorana(dcveg[,-1])

plot(ord, type = "n")

points(ord, display = "species", cex = 0.1, pch = 21, col = "black", bg = "gray")

 #ordiellipse(ord, groups$Group, display = "sites", kind = "sd", conf = 0.95, draw = "polygon")

sel <- orditorp(ord, dis = "sites", type = "n", col = "blue")
text(ord, display = "sites", cex = 0.7, col = "red")

 #sel2 <- orditorp(ord, dis = "spec", cex = 0.7, col = "red")

```

```{r fileindex, include=FALSE}

## ----------------- CREATE INDEX FILE  ----------------------

library(dplyr)

filepath <- "C:\\Users\\Meg\\OneDrive - University of Nebraska-Lincoln\\R practice\\DCVegMon\\data\\"

file.list <- list.files(filepath, pattern = '*.xlsx', full.names = TRUE) 

extract_data2 <- function(x)
{
relevecode <- read_xlsx(x, sheet = "Woody", range = "C2:C2", col_names = FALSE, na="")  

## Save this rabbit hole for later. Blank cells don't exist in Excel, and R will only provide the Index File for files that have all cells. Not sure at which step this is happening - in read or during the bind? 
#relevename <- read_xlsx(x, sheet = "Woody", range ="B2:B2", col_names = FALSE, na="")
#relevedate <- read_xlsx(x, sheet = "Woody", range = "B5:B5", col_names = FALSE, na="")

df <- x 

df <- bind_cols(df, relevecode)#, relevename, relevedate)

return(df)
}

IndexFile <- data.frame("file"= character(), "releve code" = character(), "releve name" = character(), "releve date" = character())

for (x in file.list){
  single_file2 <- extract_data2(x)
  IndexFile <- rbind(IndexFile, single_file2)
}

## Compare the Index File to the releve codes that have been imported from excel. Report missing files
All <-unique(all_df$releveCode)
Index <- unique(IndexFile$...2)
Missing <- Index[!(Index %in% All)]

print(Missing)
#print(IndexFile)

```

## OH NOS!

The following releves have issues. Check for blank/missing data in the observations worksheets.

```{r uhoh , echo=FALSE}
print(Missing)
```

## Project To Do:

-   xread 2020 and 2021 releve data
-   xread GRG reference (ordination folder)
-   xTable: Releve, Species, Midpoint % Cover (column M), year
-   xFor our data: Need to add name column to each - in C2 of Woody tab \*xadd column (2020, 2021)
-   xFor GRG data: the name will be in the NPC Code or whatever's in A1
-   xPull in the Compiled Releve Data for R, sheet1 - add year column
-   xCreate matrix
-   Groups Table - add later?
-   low priority: quality control for missing documents
-   add the groups back into the plot
-   remove labels, change symbology
-   Appsheet for data collection for future
-   double check that the data format
-   idea: k means clustering analysis? bootstrap -\> 95% confidence
-   GRG Height Classes - for future analysis work and also for the FQI. Not required for ordination script
