---
title: "Vegetation Monitoring Analysis"
author: "Manhatton"
date: "2/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

##install packages
#install.packages('skimr', repos = "http://cran.us.r-project.org")
#install.packages('kableExtra', repos = "http://cran.us.r-project.org")
#install.packages('janitor', repos = "http://cran.us.r-project.org")

## attach libraries
library(tidyverse)
library(readxl)
library(janitor)
library(dplyr)
library(here) # this was being buggy in the extract function so I removed it
library(skimr)  # not used so far
library(kableExtra)  #not used so far
```


``` {r DCrelevedata, include = false}

## Change this to the filepath for the data directory

directory <- "C:\\Users\\Meg\\OneDrive - University of Nebraska-Lincoln\\R practice\\DCVegMon\\data\\"

##Find all releve xlsx within directory

file.list <- list.files(directory, pattern = '*.xlsx', full.names = TRUE)

## Extract data from woody, forbs and grasses worksheets

 extract_data <- function(x)
{
woody <- read_xlsx(x, sheet = "Woody", skip = 8)%>%
  janitor::clean_names() %>%
  janitor::remove_empty(c("rows"))%>%
  mutate(rarity_status = as.character(rarity_status)) %>%
  select("scientific_name", "common_name", "x3" , "physiognamy", "mn_nativity", "mnwi", "rarity_status" = , "id" , "cc" , "cc_range", "x12" , "midpoint_cc", "p", "p_c")

forbs <- read_xlsx(x, sheet = "Forbs", skip = 8)%>%
 janitor::clean_names() %>%
 janitor::remove_empty(c("rows"))%>%
 mutate(rarity_status = as.character(rarity_status))  %>%
 select("scientific_name", "common_name", "x3" , "physiognamy", "mn_nativity", "mnwi", "rarity_status" = , "id" , "cc" , "cc_range", "x12" , "midpoint_cc", "p", "p_c")
  
grasses <- read_xlsx(x, sheet = "Grasses", skip = 8)%>%
  janitor::clean_names() %>%
  janitor::remove_empty(c("rows"))%>%
  mutate(rarity_status = as.character(rarity_status)) %>%
  select("scientific_name", "common_name", "x3" , "physiognamy", "mn_nativity", "mnwi", "rarity_status" = , "id" , "cc" , "cc_range", "x12" , "midpoint_cc", "p", "p_c")

relevecode <- read_xlsx(x, sheet = "Woody", range = "C2:C2", col_names = FALSE) #grabs the relevecode from C2 of Woody sheet

 data_frame <- (bind_rows(woody, forbs, grasses)) #combine the woody, forbs, and grasses data rows

 data_frame['releveCode'] <- list(relevecode[[1,1]]) #assign value from woody cell C2 to new column in combined dataframe (indicates releve code)
 
 ##getting the year out of the releve code
 n_last <- 4
 data_frame ['releveyear'] <- substr(relevecode, nchar(relevecode)-n_last + 1, nchar(relevecode) )
  
  return(data_frame)
}

 ##create empty dataframe with all the required columns
 
all_df <- data.frame("scientific_name"=character(), "common_name"=character(), "x3" =numeric(), "physiognamy"=character(), "mn_nativity" = character(), "mnwi"= character(), "rarity_status" = character(), "id" = character(), "cc" = numeric(), "cc_range" = character(), "x12" = numeric(), "midpoint_cc"= numeric(), "p" = numeric(), "p_c" = numeric(), releveCode = character(), releveyear = numeric()) 

##loop through all files in the folder 

for (x in file.list){
  print(x)
  single_df <- extract_data(x)
  all_df <- rbind(all_df, single_df)
}

##removes all rows with greater than 10 "NA" cells 

hi <- apply(all_df, 1, function(x) sum(is.na(x)))
all_df <- all_df[hi<10,]

```

``` {r DC_Old_Data, include = false}
## I was going to pull in the data that was processed in 2019 for the ordination project, but can we just put the original "Analyzed Data" files in with the 2020 and 2021 spreadsheets and use the code in DCrelevedata chunk? That would be easier :) Unless the format of the original .xls files has changed...

```

``` {r ReferenceSiteData, include=FALSE}

## This is the code to load in the GRG Reference Site Data ---------------------

## change this to the correct directory
refdir <- "C:\\Users\\Meg\\OneDrive - University of Nebraska-Lincoln\\R practice\\DCVegMon\\data\\GRG Reference"

file.list <- list.files(refdir, pattern = '*.xlsx', full.names = TRUE)

ref_extract <- function(y)
{ 
  refdata <- read_xlsx(y, sheet= 1, skip = 1, col_names = TRUE) %>%
    janitor::remove_empty(c("rows"))%>%
    select("Genus", "Species", "3Freq" , "4Abund", "5Index")
  
  refcode <- read_xlsx(y, sheet = 1, range = "A1:A1", col_names = FALSE)
  refcode <- list(refcode[[1,1]])
  
  df<- bind_cols(refdata, refcode)
  
  return(df)
}

all_ref <- data.frame("genus" = character(), "species" = character(), "freq" = numeric() , "abund" = numeric(), "index"= numeric())

for (y in file.list){
  print(y)
  single_ref <- ref_extract(y)
  all_ref <- rbind(all_ref, single_ref)
}

##removes all rows with greater than 4 "NA" cells 

empties <- apply(all_ref, 1, function(x) sum(is.na(x)))
all_ref <- all_ref[empties<4,]

```

``` {r ordination}
##create new dataframe with just the scientific name, midpoint % cover, releve 

#dc_ord_df <- all_df() 

```

```{r fileindex, include=FALSE}

## ----------------- CREATE INDEX FILE  ----------------------

library(dplyr)

filepath <- "C:\\Users\\Meg\\OneDrive - University of Nebraska-Lincoln\\R practice\\DCVegMon\\data\\"

file.list <- list.files(filepath, pattern = '*.xlsx', full.names = TRUE) 

extract_data2 <- function(x)
{
relevecode <- read_xlsx(x, sheet = "Woody", range = "C2:C2", col_names = FALSE, na="")  

## Save this rabbit hole for later. Blank cells don't exist in Excel, and R will only provide the Index File for files that have all cells. Not sure at which step this is happening - in read or during the bind? 
#relevename <- read_xlsx(x, sheet = "Woody", range ="B2:B2", col_names = FALSE, na="")
#relevedate <- read_xlsx(x, sheet = "Woody", range = "B5:B5", col_names = FALSE, na="")

df <- x 

df <- bind_cols(df, relevecode)#, relevename, relevedate)

return(df)
}

IndexFile <- data.frame("file"= character(), "releve code" = character(), "releve name" = character(), "releve date" = character())

for (x in file.list){
  single_file2 <- extract_data2(x)
  IndexFile <- rbind(IndexFile, single_file2)
}

## Compare the Index File to the releve codes that have been imported from excel. Report missing files
All <-unique(all_df$releveCode)
Index <- unique(IndexFile$...2)
Missing <- Index[!(Index %in% All)]

print(Missing)
#print(IndexFile)

```

## OH NOS!
The following releves have issues. Check for blank/missing data in the observations worksheets.

```{r uhoh , echo=false}
print(Missing)
```

## Project To Do:

*xread 2020 and 2021 releve data
*xread GRG reference (ordination folder)
*-Table: Releve, Species, Midpoint % Cover (column M), year
*xFor our data: Need to add name column to each - in C2 of Woody tab *xadd column (2020, 2021)
*xFor GRG data: the name will be in the NPC Code or whatevers in A1
*-Pull in the Compiled Releve Data for R, sheet1 - add year column
    +oy, see note in DC_Old_Data chunk
*-Create matrix


```{r df}


```

